# Machine Learning Course Project

This project uses the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants in the [Weight Lifting Exercise Dataset](http://groupware.les.inf.puc-rio.br/har) to predict whether a barbell lift is being performed correctly.

## Data Cleansing

We first download the [Training Data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv) and [Test Data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv) and read them into R.

```{r}
trainPml <- read.csv("pml-training.csv", na.strings=c("NA",""))
testPml <- read.csv("pml-testing.csv", na.strings=c("NA",""))
```

Checking the datasets for NA values revealed a large number of variables where almost all values are missing.  These variables are removed since they are unlikely to add any value to the prediction model.

```{r}
table(colSums(is.na(trainPml)))
table(colSums(is.na(testPml)))
missingVar <- colSums(is.na(trainPml))==19216
trainPml <- trainPml[!missingVar]
testPml <- testPml[!missingVar]
```

Similarly, the first 7 columns also appear to be contain non-predictive information, and they are thus removed from the training and test sets.

```{r}
str(trainPml[1:7])
trainPml <- trainPml[-c(1:7)]
testPml <- testPml[-c(1:7)]
```

## Model building

```{r}
library(caret)
library(randomForest)
```

The training data is further partitioned into training and test sets for the purpose of model building.

```{r}
set.seed(12345)
inTrain <- createDataPartition(y=trainPml$classe, p=0.6, list=FALSE)
training <- trainPml[inTrain,]
testing <- trainPml[-inTrain,]
```

The Random Forest technique is used to build a simple predictive model.

```{r}
model <- randomForest(classe~., data=training)
```

We compute the out of sample error on the test set as follows.

```{r}
confusionMatrix(predict(model, testing), testing$classe)
```

The expected out of sample error is thus `0.7%`.

## Predictions

We make predictions for the test set and write the predictions to individual text files as required.

```{r}
answers <- as.character(predict(model, testPml))
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```

The model obtains a prediction of 20/20 when submitted to Coursera.